Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms

### 2018-02-28 paper/paper.tex
Little tweaks

\begin{abstract}
    We discuss our experiences adapting three recent algorithms for maximum common (connected)
    subgraph problems to exploit multi-core parallelism. These algorithms are heavily irregular and
    very sensitive to search order changes, and so do not easily lend themselves to parallel search.
    Nonetheless, our results show that each algorithm can be parallelised successfully, with the
    threaded algorithms we create being clearly better than the sequential ones. We then look in
    more detail at the results, and discuss how speedups should be measured for this kind of algorithm.
    Because of the difficulty in quantifying an average speedup when so-called anomalous
    speedups (superlinear and sublinear) are common, we propose a new measure called \emph{aggregate
    speedup}.
\end{abstract}

@@

Two approaches exist for ensuring connectedness: either a conventional global constraint and
propagator can be used \cite{DBLP:conf/cp/McCreeshNPS16}, or a special branching rule can be used
to enforce connectedness during search \cite{DBLP:conf/mco/VismaraV08}. The two techniques are
broadly comparable performance-wise \cite{DBLP:conf/cp/McCreeshNPS16}, but the branching rule is
much simpler to implement.

@@

    \item[Vertex labelled] undirected instances, by selecting the first ten members of each
        parameter class (and so graphs have up to 100 vertices each), and applying the 33\%
        labelling scheme for vertices only. This gives 8,140 instances.

    \item[Both labelled, directed] instances, by selecting the first ten members of each parameter
        class, and applying the 33\% labelling scheme to both vertices and edges. Again, this gives
        8,140 instances.

@@

Parallel tree-search has a long history \cite{o:BaderHC05}. Of particular interest to us are
so-called \emph{anomalies}
\cite{DBLP:journals/cacm/LaiS84,DBLP:journals/tc/LiW86,DBLP:conf/irregular/BruinKT95}: because we
are not performing a fixed amount of work, we should have no expectation of a linear speedup, and
instead we could see a sublinear speedup (much less than $n$ from $n$ processors, if speculative
work turns out to be wasted) or a superlinear speedup (much more than $n$ from $n$ processors, if a
strong incumbent is found more quickly). An absolute slowdown (a speedup much less than 1) is also
possible under some models.

@@

We will use this mechanism for our experiments.  The clique-based maximum common subgraph algorithm
effectively differs only in the preprocessing stage, and the clique-inspired connected algorithm
described by McCreesh et al.\ \cite{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that it may be parallelised in
exactly the same way. Based upon preliminary experiments, we use a splitting depth limit of five
rather than the original three, since maximum common subgraph instances appear to give even more
irregular search trees than normal clique problem instances.

@@

In \cref{figure:kdownscatters} we show per-instance and aggregate speedups for the $k{\downarrow}$
algorithm. On unlabelled instances, we see a range of speedups between slightly less than one and
ten, with an aggregate speedup of seven. These results are not as good as with the clique algorithm.
Profiling suggests memory allocation problems: although the amount of work done would suggest good
parallelism, the time taken to perform each domain copy operation increases as the number of threads
increases. Unlike the clique algorithm, which has very small, cache-friendly data structures which
are modified in-place, the state for the $k{\downarrow}$ algorithm is large and much of the runtime
is spent copying data structures. (Our hardware is a dual multi-core processor configuration, and
each core has its own low-level cache, but memory bandwidth is shared. Interestingly, on older Xeon
E5 v2 systems, this problem is much more pronounced.)

>>>>

\begin{abstract}
    We discuss our experiences adapting three recent algorithms for maximum common (connected)
    subgraph problems to exploit multi-core parallelism. These algorithms do not easily lend
    themselves to parallel search, as the search trees are extremely irregular, making balanced work
    distribution hard, and runtimes are very sensitive to value-ordering heuristic behaviour.
    Nonetheless, our results show that each algorithm can be parallelised successfully, with the
    threaded algorithms we create being clearly better than the sequential ones. We then look in
    more detail at the results, and discuss how speedups should be measured for this kind of
    algorithm.  Because of the difficulty in quantifying an average speedup when so-called anomalous
    speedups (superlinear and sublinear) are common, we propose a new measure called \emph{aggregate
    speedup}.
\end{abstract}

@@

Two approaches exist for ensuring connectedness: either a conventional global constraint and
propagator can be used \cite{DBLP:conf/cp/McCreeshNPS16}, or a special branching rule can
enforce connectedness during search \cite{DBLP:conf/mco/VismaraV08}. The two techniques are
broadly comparable performance-wise \cite{DBLP:conf/cp/McCreeshNPS16}, but the branching rule is
simpler to implement.

@@

    \item[Vertex labelled] undirected instances, by selecting the first ten members of each
        parameter class (and so graphs have up to 100 vertices each), using the 33\%
        labelling scheme \cite{DBLP:journals/prl/SantoFSV03} for vertices only. This gives 8,140 instances.

    \item[Both labelled, directed] instances, by selecting the first ten members of each parameter
        class, and applying the 33\% labelling scheme \cite{DBLP:journals/prl/SantoFSV03} to both
        vertices and edges. Again, this gives 8,140 instances.

@@

Parallel tree-search has a long history \cite{o:BaderHC05}. Of particular interest to us are
so-called \emph{anomalies}
\cite{DBLP:journals/cacm/LaiS84,DBLP:journals/tc/LiW86,DBLP:conf/irregular/BruinKT95}: because we
are not performing a fixed amount of work, we should have no expectation of a linear speedup, and
instead we could see a sublinear speedup (much less than $n$ from $n$ processors, if speculative
work turns out to be wasted) or a superlinear speedup (much more than $n$ from $n$ processors, if a
strong incumbent is found more quickly). An absolute slowdown (a speedup much less than 1) is also
possible when using some parallelisation techniques.

@@

We will use this mechanism for our experiments.  The clique-based maximum common subgraph algorithm
effectively differs only in the preprocessing stage, and the clique-inspired connected algorithm
described by McCreesh et al.\ \cite{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that it may be parallelised in
exactly the same way. Based upon preliminary experiments, we set the mechanism's splitting depth
limit parameter to be five rather than the original three, since maximum common subgraph instances
appear to give even more irregular search trees than normal clique problem instances.

@@

In \cref{figure:kdownscatters} we show per-instance and aggregate speedups for the $k{\downarrow}$
algorithm. On unlabelled instances, we see a range of speedups between 0.9 and ten, with an
aggregate speedup of seven. These results are not as good as with the clique algorithm.  Profiling
suggests memory allocation problems: although the amount of work done would suggest good
parallelism, the time taken to perform each domain copy operation increases as the number of threads
increases. Unlike the clique algorithm, which has very small, cache-friendly data structures which
are modified in-place, the state for the $k{\downarrow}$ algorithm is large and much of the runtime
is spent copying data structures. (Our hardware is a dual multi-core processor configuration, and
each core has its own low-level cache, but memory bandwidth is shared. Interestingly, on older Xeon
E5 v2 systems, this problem is much more pronounced.)
